---
title: 'Supplementary Material for "Morph-dependent effect of nematode infection on host movement in the land snail *Cepaea nemoralis* (Mollusca, Gastropoda)"'
author: "Maxime Dahirel, Marine Proux, Armelle Ansart, Claudia Gérard"
date:
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE,
  message = FALSE, eval = TRUE, dev = "cairo_pdf"
)
```

```{r packages-loading}
library(matrixStats) # CRAN v0.61.0

library(cmdstanr) # [github::stan-dev/cmdstanr] v0.4.0.9000
library(brms) # CRAN v2.16.3

library(tidyverse) # CRAN v1.3.1

library(bayesplot) # CRAN v1.8.1
library(tidybayes) # CRAN v3.0.2

library(patchwork) # CRAN v1.1.1

library(here) # CRAN v1.0.1

options(mc.cores = 4) ## to reduce/increase depending on cores available on your machine
```


# S1 - detailed math description of both models

## Size model


## Multivariate model of parasites abundance and snail behaviours


# S2 - descriptive stats about parasites 

```{r data-load}
raw_parasite <- read_csv(here("data", "cepaea_2019_parasite.csv"))
raw_behaviour <- read_csv(here("data", "cepaea_2019_behaviour.csv"))

data_parasite <- raw_parasite %>%
  mutate(population = fct_relevel(factor(population), "shaded", after = Inf)) %>% # factor reordering
  mutate(Nematode_live_all = Nematode_0_1_free + Nematode_1_3_free +
           Nematode_3_5_free + Nematode_5_plus_free + Nematode_1_3_lung + 
           Nematode_3_5_lung + Nematode_5_plus_lung,
         Nematode_encaps_all = Nematode_white_encaps + Nematode_dark_encaps,
         Acari_live_all = Acari_kidney + Acari_digestive
         ) %>% 
  mutate(session = 1)

data <- raw_behaviour %>%
  mutate(population = fct_relevel(factor(population), "shaded", after = Inf)) %>%
  left_join(data_parasite) %>%
  filter(dead == 0)

data <- data %>%
  # we then standardize size and response variables
  mutate(
    scale_size = scale(shell_diameter)[, 1],
    scale_food = scale(Food_intake)[, 1]
  ) %>%
  mutate(
    mean_size = mean(shell_diameter, na.rm = TRUE), ## useful for back-transformations for plots
    sd_size = sd(shell_diameter, na.rm = TRUE)
  ) %>%
  mutate(
    mean_food = mean(Food_intake, na.rm = TRUE),
    sd_food = sd(Food_intake, na.rm = TRUE)
  )
```

We present here information about each parasite prevalence (% of potential hosts infected) and infection intensity
(number of parasites per *actually infected host*). For trematode and nematode data, the provided 95% intervals are the Highest Density intervals based on binomial GLMs (prevalence) and negative binomial GLMs (intensity) with morph, population and morph *population fixed effects. The intensity models are fitted on intensity - 1, to account for the zero exclusion. Priors are as in S1, with the addition of a halfnormal(0-1) prior on the inverse of the negative binomial shape parameter, per REF. Note that we do not provide 95% intervals for Acari, given rarity of infection. and also not for encapsulated parasites, for reasons presented at the end of the main text methods.
The individual that died is excluded from all tables below and harboured `r subset(data_parasite$Nematode_live_all,data_parasite$dead==1)[1]` live nematodes, `r subset(data_parasite$Trematode_kidney,data_parasite$dead==1)[1]` live trematodes, `r subset(data_parasite$Acari_live_all,data_parasite$dead==1)[1]` live mites, had encapsulated `r subset(data_parasite$Nematode_encaps_all,data_parasite$dead==1)[1]` nematodes and `r subset(data_parasite$Acari_encaps,data_parasite$dead==1)[1]` mites.



(morph and population specific prevalence, as well as min/mean/median/max CI abundance and/or intensities)

```{r}
if (file.exists(here("R_output", "suppl_model_nema_prevalence.RDS")))
  {
    mod_nemaprev <- readRDS(here("R_output", "suppl_model_nema_prevalence.RDS"))
  } else {
    
mod_nemaprev <- brm(
  bf(Nematode_live_all>0  ~  0 + band_number,
     family=bernoulli),
  data = subset(data_parasite, population!="shaded"), chains = 4, iter = 2000, warmup = 1000,
  prior = c(
    set_prior("normal(0,1)", class = "b")
  ),
  backend = "cmdstanr", seed = 42)

  saveRDS(mod_nemaprev, file = here("R_output", "suppl_model_nema_prevalence.RDS"))
  }
```


```{r}
if (file.exists(here("R_output", "suppl_model_nema_intensity.RDS")))
  {
    mod_nemaintens <- readRDS(here("R_output", "suppl_model_nema_intensity.RDS"))
  } else {
    
mod_nemaintens <- brm(
  bf(Nematode_live_all - 1  ~  0 + band_number,
  nlf(shape~1/invshape),
     invshape~1,
     family=negbinomial(link_shape = "identity")),
  data = subset(data_parasite, Nematode_live_all>0 & population!="shaded"), chains = 4, iter = 2000, warmup = 1000,
  prior = c(
    set_prior("normal(0,1)", class = "b"),
    set_prior("normal(0,1)",nlpar="invshape",lb=0)
  ),
  backend = "cmdstanr", seed = 42)  

saveRDS(mod_nemaintens, file = here("R_output", "suppl_model_nema_intensity.RDS"))
  }
```


```{r}
if (file.exists(here("R_output", "suppl_model_trema_prevalence.RDS")))
  {
    mod_tremaprev <- readRDS(here("R_output", "suppl_model_trema_prevalence.RDS"))
  } else {
    
mod_tremaprev <- brm(
  bf(Trematode_kidney>0  ~  0 + band_number + band_number : population,
     family=bernoulli),
  data = data_parasite, chains = 4, iter = 2000, warmup = 1000,
  prior = c(
    set_prior("normal(0,1)", class = "b")
  ),
  backend = "cmdstanr", seed = 42)

  saveRDS(mod_tremaprev, file = here("R_output", "suppl_model_trema_prevalence.RDS"))
  }
```


```{r}
if (file.exists(here("R_output", "suppl_model_trema_intensity.RDS")))
  {
    mod_tremaintens <- readRDS(here("R_output", "suppl_model_trema_intensity.RDS"))
  } else {
    
mod_tremaintens <- brm(
  bf(Trematode_kidney - 1  ~  0 + band_number + band_number : population,
  nlf(shape~1/invshape),
     invshape~1,
     family=negbinomial(link_shape = "identity")),
  data = subset(data_parasite, Trematode_kidney>0), chains = 4, iter = 2000, warmup = 1000,
  prior = c(
    set_prior("normal(0,1)", class = "b"),
    set_prior("normal(0,1)",nlpar="invshape",lb=0)
  ),
  backend = "cmdstanr", seed = 42)

saveRDS(mod_tremaintens, file = here("R_output", "suppl_model_trema_intensity.RDS"))
  }
```

```{r}
nema_obs <- data_parasite %>%
  filter(dead==0) %>% 
  group_by(population, band_number) %>%
  summarise(n = length(Nematode_live_all),
            n_infected = length(subset(Nematode_live_all,Nematode_live_all>0)),
            prevalence = mean(Nematode_live_all > 0)%>% round(digits=2),
            mean_intensity = mean(subset(Nematode_live_all,Nematode_live_all>0))%>% round(digits=2),
            min_intensity = min(subset(Nematode_live_all,Nematode_live_all>0))%>% round(digits=2),
            max_intensity = max(subset(Nematode_live_all,Nematode_live_all>0))%>% round(digits=2)) %>% 
  ungroup() %>% 
  mutate(mean_intensity=round(mean_intensity,2) %>% replace_na("--"),
         max_intensity=replace(max_intensity,which(max_intensity<0),"--"),
         min_intensity=replace(min_intensity,which(min_intensity>1000),"--"),
         )


nema_pred_prev <- data_parasite %>% 
  filter(population!="shaded") %>% 
  select(band_number,population) %>% 
  distinct() %>% 
  add_epred_draws(mod_nemaprev) %>% 
  mutate(prevalence=.epred) %>% 
  group_by(band_number,population) %>% 
  mean_hdi(prevalence) %>% 
  group_by(band_number,population) %>% 
  summarise(.lower=min(.lower),.upper=max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>% 
  mutate(prevalence_interval = paste0("[", round(.lower, 2), "; ", round(.upper, 2), "]", sep = ""))%>% 
  select(band_number,population,prevalence_interval)


nema_pred_intens <- data_parasite %>% 
  filter(population!="shaded") %>% 
  select(band_number,population) %>% 
  distinct() %>% 
  add_epred_draws(mod_nemaintens) %>% 
  mutate(intensity=.epred+1) %>% 
  group_by(band_number,population) %>% 
  mean_hdi(intensity) %>% 
  group_by(band_number,population) %>% 
  summarise(.lower=min(.lower),.upper=max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>% 
  mutate(intensity_interval = paste0("[", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) %>% 
  select(band_number,population,intensity_interval)


table_nema=left_join(nema_obs,nema_pred_prev) %>%
  left_join(nema_pred_intens) %>%
  mutate(prevalence_interval=replace_na(prevalence_interval,"--"),
         intensity_interval=replace_na(intensity_interval,"--")) %>% 
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                             `0 bands` ="0B",
                             `3 bands` = "3B",
                             `5 bands` = "5B")) %>%  
  select(population,band_number,
         n,n_infected,prevalence,prevalence_interval,
         mean_intensity, intensity_interval,
         min_intensity, max_intensity)

```




```{r}
trema_obs <- data_parasite %>%
  filter(dead==0) %>% 
  group_by(population, band_number) %>%
  summarise(n = length(Trematode_kidney),
            n_infected = length(subset(Trematode_kidney,Trematode_kidney>0)),
            prevalence = mean(Trematode_kidney > 0)%>% round(digits=2),
            mean_intensity = mean(subset(Trematode_kidney,Trematode_kidney>0))%>% round(digits=2),
            min_intensity = min(subset(Trematode_kidney,Trematode_kidney>0))%>% round(digits=2),
            max_intensity = max(subset(Trematode_kidney,Trematode_kidney>0))%>% round(digits=2)) %>% 
  ungroup()


trema_pred_prev <- data_parasite %>% 
  select(band_number,population) %>% 
  distinct() %>% 
  add_epred_draws(mod_tremaprev) %>% 
  mutate(prevalence=.epred) %>% 
  group_by(band_number,population) %>% 
  mean_hdi(prevalence) %>% 
  group_by(band_number,population) %>% 
  summarise(.lower=min(.lower),.upper=max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>% 
  mutate(prevalence_interval = paste0("[", round(.lower, 2), "; ", round(.upper, 2), "]", sep = ""))%>% 
  select(band_number,population,prevalence_interval)


trema_pred_intens <- data_parasite %>% 
  select(band_number,population) %>% 
  distinct() %>% 
  add_epred_draws(mod_tremaintens) %>% 
  mutate(intensity=.epred+1) %>% 
  group_by(band_number,population) %>% 
  mean_hdi(intensity) %>% 
  group_by(band_number,population) %>% 
  summarise(.lower=min(.lower),.upper=max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>% 
  mutate(intensity_interval = paste0("[", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) %>% 
  select(band_number,population,intensity_interval)


table_trema = left_join(trema_obs,trema_pred_prev) %>%
  left_join(trema_pred_intens) %>%
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                             `0 bands` ="0B",
                             `3 bands` = "3B",
                             `5 bands` = "5B")) %>%  
  select(population,band_number,
         n,n_infected,prevalence,prevalence_interval,
         mean_intensity, intensity_interval,
         min_intensity, max_intensity)

```


```{r}
Acari_obs <- data_parasite %>%
  filter(dead==0) %>% 
  group_by(population, band_number) %>%
  summarise(n = length(Acari_live_all),
            n_infected = length(subset(Acari_live_all,Acari_live_all>0)),
            prevalence = mean(Acari_live_all > 0) %>% round(digits=2),
            mean_intensity = mean(subset(Acari_live_all,Acari_live_all>0))%>% round(digits=2),
            min_intensity = min(subset(Acari_live_all,Acari_live_all>0)),
            max_intensity = max(subset(Acari_live_all,Acari_live_all>0))) %>% 
  ungroup()%>% 
  mutate(mean_intensity=round(mean_intensity,2) %>% replace_na("--"),
         max_intensity=replace(max_intensity,which(max_intensity<0),"--"),
         min_intensity=replace(min_intensity,which(min_intensity>1000),"--"),
         ) %>%
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                             `0 bands` ="0B",
                             `3 bands` = "3B",
                             `5 bands` = "5B"))
```


```{r}
encaps_obs <- data_parasite %>%
  filter(dead==0) %>% 
  group_by(population, band_number) %>%
  summarise(n = length(Acari_encaps),
            p_encaps_nema = mean(Nematode_encaps_all > 0) %>% round(digits=2),
            p_encaps_mite = mean(Acari_encaps > 0) %>% round(digits=2)) %>% 
  ungroup()%>% 
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                             `0 bands` ="0B",
                             `3 bands` = "3B",
                             `5 bands` = "5B"))
```



table x nematode
```{r}
knitr::kable(table_nema %>% 
               arrange(population,band_number) %>% 
               select(`shell morph` = band_number,n,
                      prevalence,`95% interval`=prevalence_interval), booktabs = T) %>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

table x trematode
```{r}
knitr::kable(table_trema %>% 
               arrange(population,band_number) %>% 
               select(`shell morph`=band_number,n,
                      prevalence,`95% interval`=prevalence_interval), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

Table x acari
```{r}
knitr::kable(Acari_obs %>% 
               arrange(population,band_number) %>%  select(`shell morph`=band_number,n,prevalence), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

table x nematode
```{r}
knitr::kable(table_nema %>% 
               arrange(population,band_number) %>% 
               select(`shell morph`=band_number,
                                   `n(infected)`=n_infected,
                                   `mean intensity`=mean_intensity, 
                                   `95% interval`=intensity_interval,
                                   min = min_intensity, max = max_intensity), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

table x trematode
```{r}
knitr::kable(table_trema %>% 
               arrange(population,band_number) %>% 
               select(`shell morph`=band_number,
                                   `n(infected)`=n_infected,
                                   `mean intensity`=mean_intensity, 
                                   `95% interval`=intensity_interval,
                                   min = min_intensity, max = max_intensity), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

Table x acari
```{r}
knitr::kable(Acari_obs %>% 
               arrange(population,band_number) %>% 
               select(`shell morph`=band_number,
                                   `n(infected)`=n_infected,
                                   `mean intensity`=mean_intensity, 
                                   min = min_intensity, max = max_intensity), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```

Table x encaps
```{r}
knitr::kable(encaps_obs %>% 
               arrange(population,band_number) %>%  select(`shell morph`=band_number,
                                                           n,`proportion with encapsulated nematodes`=p_encaps_nema,
                                                           `proportion with encapsulated mites`=p_encaps_mite), booktabs = T)%>%
  kableExtra::pack_rows("Open habitat", 1,3) %>%
  kableExtra::pack_rows("Shaded habitat", 4,6)
```


# S3 - pairwise posterior comparisons for each trait

```{r loading-model}
mod_multi <- readRDS(here("R_output","model_multi.RDS"))
```


```{r}
preds_food <- data %>%
  select(population, band_number, mean_food, sd_food) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1) %>%
  add_epred_draws(mod_multi, re_formula = NA, resp = "scalefood") %>%
  mutate(.epred = .epred * sd_food + mean_food) %>%
  ungroup() %>%
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                           `0 bands` = "0B",
                           `3 bands` = "3B",
                           `5 bands` = "5B"))

preds_mvt <- data %>%
  select(population, band_number) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1, is_active = TRUE) %>%
  add_epred_draws(mod_multi, re_formula = NA, resp = "scalemvt") %>%
  ungroup() %>%
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                           `0 bands` = "0B",
                           `3 bands` = "3B",
                           `5 bands` = "5B"))

preds_trematode <- data %>%
  select(band_number, population) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1) %>%
  add_linpred_draws(mod_multi, re_formula = NA, resp = "Trematodekidney") %>%
  ungroup() %>%
  mutate(population = fct_recode(population,
    `shaded habitat` = "shaded",
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                           `0 bands` = "0B",
                           `3 bands` = "3B",
                           `5 bands` = "5B"))

preds_nematode <- data %>%
  select(band_number) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1, population = "sun_exposed") %>%  ## predictions are made at the average size
  add_linpred_draws(mod_multi, re_formula = NA, resp = "Nematodeliveall") %>%
  ungroup() %>%
  mutate(population = fct_recode(population,
    `open habitat` = "sun_exposed"
  ),
  band_number = fct_recode(band_number,
                           `0 bands` = "0B",
                           `3 bands` = "3B",
                           `5 bands` = "5B"))

```

```{r}

preds_nematode %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.linpred, by = group) %>%
  ggplot()+
  stat_halfeye(aes(.linpred,group),.width=c(0.001,0.95),point_interval="mean_hdi") +
  geom_vline(xintercept=0,lty=2)+
  scale_y_discrete("pair compared")+
  scale_x_continuous("posterior predicted difference (latent log scale)")+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")

preds_trematode %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.linpred, by = group) %>%
  ggplot()+
  stat_halfeye(aes(.linpred,group),.width=c(0.001,0.95),point_interval="mean_hdi") +
  geom_vline(xintercept=0,lty=2)+
  scale_y_discrete("pair compared")+
  scale_x_continuous("posterior predicted difference (latent log scale)")+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")

preds_mvt %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.epred, by = group) %>%
  ggplot()+
  stat_halfeye(aes(.epred,group),.width=c(0.001,0.95),point_interval="mean_hdi") +
  geom_vline(xintercept=0,lty=2)+
  scale_y_discrete("pair compared")+
  scale_x_continuous("posterior predicted difference")+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")

preds_food %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.epred, by = group) %>%
  ggplot()+
  stat_halfeye(aes(.epred,group),.width=c(0.001,0.95),point_interval="mean_hdi") +
  geom_vline(xintercept=0,lty=2)+
  scale_y_discrete("pair compared")+
  scale_x_continuous("posterior predicted difference")+
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```

# S4 -  Posterior proportion of total variance associated with fixed effects vs. random effects in movement and food intake

Both movement behaviour and food intake were observed twice, which allows us to partition variance into among-individual and within-individual components. We show that although in both cases, within-individual/"residual" variation is the dominant component, there is a non-negligible among-individual variance component: both traits are repeatable (**Supporting Information Figure SX**).

```{r fig-varpart}
## the variables used to fit models are scaled
## but that's not a problem since we are plotting **relative** variance components here

## in each case there are four variance components: 
## the fixed effects, the individual-level random effects, 
## the experimental session random effects, and the residual within-individual variation


## here for the movement model
varcomps_mvt <- tibble(
  VF = rowVars(posterior_linpred(mod_multi, re_formula = NA, resp = "scalemvt")),
  VI = VarCorr(mod_multi, summary = FALSE)$fullID$sd[, "scalemvt_Intercept"]^2,
  Vseries = VarCorr(mod_multi, summary = FALSE)$series$sd[, "scalemvt_Intercept"]^2,
  VR = as_tibble(as_draws_df(mod_multi, variable="sigma_scalemvt"))$sigma_scalemvt^2
) %>%
  mutate(
    trait = "mvt",
    VP = VF + VI + Vseries + VR
  )

## and here for the food intake model
varcomps_food <- tibble(
  VF = rowVars(posterior_linpred(mod_multi, re_formula = NA, resp = "scalefood")),
  VI = VarCorr(mod_multi, summary = FALSE)$fullID$sd[, "scalefood_Intercept"]^2,
  Vseries = VarCorr(mod_multi, summary = FALSE)$series$sd[, "scalefood_Intercept"]^2,
  VR = as_tibble(as_draws_df(mod_multi, variable="sigma_scalefood"))$sigma_scalefood^2
) %>%
  mutate(
    trait = "food",
    VP = VF + VI + Vseries + VR
  )

## we merge them and do some rearranging/ renaming
varcomps <- rbind(varcomps_food, varcomps_mvt) %>%
  group_by(trait) %>%
  pivot_longer(cols = c(VF, VI, Vseries, VR), names_to = "varcomp", values_to = "variance") %>%
  ungroup() %>%
  mutate(varcomp = fct_recode(varcomp,
    `fixed effects` = "VF",
    individual = "VI",
    `test group` = "Vseries",
    residuals = "VR"
  )) %>%
  mutate(varcomp = fct_relevel(varcomp, rev)) %>%
  mutate(varcomp = fct_relevel(varcomp, "test group", after = 1)) %>%
  mutate(trait = fct_recode(trait,
    `Food intake` = "food",
    `Movement (1/latency to leave)` = "mvt"
  ))

## and then we plot

ggplot(varcomps) +
  stat_halfeye(aes(y = varcomp, x = 100 * variance / VP),
    point_interval = mean_hdi, .width = c(0.001, 0.95),
    normalize = "xy"
  ) +
  facet_wrap(~trait) +
  scale_x_continuous(
    name = "Variance explained (%, observed scale)", lim = c(0, 100),
    breaks = seq(from = 0, to = 100, by = 10)
  ) +
  scale_y_discrete(name = "Variance component") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```

**Supporting Information Figure SX.** Mean (points) and posteriors for the proportion of variance explained by the different variance components. See **Methods** and **Supporting Information S1** for a description of the model underlying these estimates.

# References