---
title: 'Supplementary Material for "Morph-dependent effect of nematode infection on host movement in the land snail *Cepaea nemoralis* (Mollusca, Gastropoda)"'
author: "Maxime Dahirel, Marine Proux, Armelle Ansart, Claudia Gérard"
date:
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE,
  message = FALSE, eval = TRUE, dev = "cairo_pdf"
)
## cairo is needed as device to properly render the plots with greek letters in pdf
```

```{r load-packages}

library(tidyverse)   # CRAN v1.3.1

library(here)        # CRAN v1.0.1
```

# S1 - detailed math description of both models

# S2 - descriptive stats about parasites 

(morph and population specific prevalence, as well as min/mean/median/max CI abundance and/or intensities)


# S3 - pairwise posterior comparisons for each trait

# S4 -  Posterior proportion of total variance associated with fixed effects vs. random effects, for both behavioural traits with repeated measures. 

See **Methods** and **Supplementary Material 1** for a description of the model underlying these estimates.


# References