---
title: 'Supplementary Material for "Morph-dependent effect of nematode infection on host movement in the land snail *Cepaea nemoralis* (Mollusca, Gastropoda)"'
author: "Maxime Dahirel, Marine Proux, Armelle Ansart, Claudia Gérard"
date:
output: pdf_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE,
  message = FALSE, eval = TRUE, dev = "cairo_pdf"
)
## cairo is needed as device to properly render the plots with greek letters in pdf
```

```{r load-packages}

library(tidyverse)   # CRAN v1.3.1

library(here)        # CRAN v1.0.1
```

# S1 - detailed math description of both models

# S2 - descriptive stats about parasites 

(morph and population specific prevalence, as well as min/mean/median/max CI abundance and/or intensities)




```{r}
data_parasite %>%
  group_by(population, band_number, dead) %>%
  summarise(
    preva_trema = mean(Trematode_kidney > 0),
    preva_nema = mean(Nematode_live_all > 0),
    preva_nencaps = mean(Nematode_encaps_all > 0),
    preva_acari = mean(Acari_live_all > 0),
    preva_aaencaps = mean(Acari_encaps > 0),
    abun_trema = mean(Trematode_kidney),
    abun_nema = mean(Nematode_live_all),
    abun_nencaps = mean(Nematode_encaps_all),
    abun_acari = mean(Acari_live_all),
    abun_aencaps = mean(Acari_encaps)
  )


data_parasite %>%
  group_by(population, band_number, dead) %>%
  filter(Trematode_kidney > 0) %>%
  summarise(
    N = length(Trematode_kidney),
    min_inci = min(Trematode_kidney),
    median_inci = median(Trematode_kidney),
    mean_inci = mean(Trematode_kidney),
    max_inci = max(Trematode_kidney)
  )


data_parasite %>%
  group_by(population, band_number, dead) %>%
  filter(Nematode_live_all > 0) %>%
  summarise(
    N = length(Nematode_live_all),
    min_inci = min(Nematode_live_all),
    median_inci = median(Nematode_live_all),
    mean_inci = mean(Nematode_live_all),
    max_inci = max(Nematode_live_all)
  )


data_parasite %>%
  group_by(population, band_number, dead) %>%
  filter(Nematode_encaps_all > 0) %>%
  summarise(
    N = length(Nematode_encaps_all),
    min_inci = min(Nematode_encaps_all),
    median_inci = median(Nematode_encaps_all),
    mean_inci = mean(Nematode_encaps_all),
    max_inci = max(Nematode_encaps_all)
  )

data_parasite %>%
  group_by(population, band_number, dead) %>%
  filter(Acari_live_all > 0) %>%
  summarise(
    N = length(Acari_live_all),
    min_inci = min(Acari_live_all),
    median_inci = median(Acari_live_all),
    mean_inci = mean(Acari_live_all),
    max_inci = max(Acari_live_all)
  )

data_parasite %>%
  group_by(population, band_number, dead) %>%
  filter(Acari_encaps > 0) %>%
  summarise(
    N = length(Acari_encaps),
    min_inci = min(Acari_encaps),
    median_inci = median(Acari_encaps),
    mean_inci = mean(Acari_encaps),
    max_inci = max(Acari_encaps)
  )
```


# S3 - pairwise posterior comparisons for each trait

# S4 -  Posterior proportion of total variance associated with fixed effects vs. random effects, for both behavioural traits with repeated measures. 

See **Methods** and **Supplementary Material 1** for a description of the model underlying these estimates.

<!--need to reimport, and so save, the big multivariate model-->

no need of complex stuff for repeatabilities, the two repeatable traits are gaussian here

```{r}
## expressed on scale components but OK because we plot % of variance
varcomps_mvt <- tibble(
  VF = rowVars(posterior_linpred(mod, re_formula = NA, resp = "scalemvt")),
  VI = VarCorr(mod, summary = FALSE)$fullID$sd[, "scalemvt_Intercept"]^2,
  Vseries = VarCorr(mod, summary = FALSE)$series$sd[, "scalemvt_Intercept"]^2,
  VR = (posterior_samples(mod, pars = "sigma_scalemvt")[, 1])^2
) %>%
  mutate(
    trait = "mvt",
    VP = VF + VI + Vseries + VR
  )

varcomps_food <- tibble(
  VF = rowVars(posterior_linpred(mod, re_formula = NA, resp = "scalefood")),
  VI = VarCorr(mod, summary = FALSE)$fullID$sd[, "scalefood_Intercept"]^2,
  Vseries = VarCorr(mod, summary = FALSE)$series$sd[, "scalefood_Intercept"]^2,
  VR = (posterior_samples(mod, pars = "sigma_scalefood")[, 1])^2
) %>%
  mutate(
    trait = "food",
    VP = VF + VI + Vseries + VR
  )

rbind(varcomps_food, varcomps_mvt) %>%
  group_by(trait) %>%
  pivot_longer(cols = c(VF, VI, Vseries, VR), names_to = "varcomp", values_to = "variance") %>%
  ungroup() %>%
  mutate(varcomp = fct_recode(varcomp,
    `fixed effects` = "VF",
    individual = "VI",
    `test group` = "Vseries",
    residuals = "VR"
  )) %>%
  mutate(varcomp = fct_relevel(varcomp, rev)) %>%
  mutate(varcomp = fct_relevel(varcomp, "test group", after = 1)) %>%
  mutate(trait = fct_recode(trait,
    `Food intake` = "food",
    `Movement speed` = "mvt"
  )) %>%
  ggplot() +
  stat_halfeye(aes(y = varcomp, x = 100 * variance / VP),
    point_interval = mean_hdi, .width = c(0.001, 0.95),
    normalize = "xy"
  ) +
  facet_wrap(~trait) +
  scale_x_continuous(
    name = "Variance explained (%, observed scale)", lim = c(0, 100),
    breaks = seq(from = 0, to = 100, by = 10)
  ) +
  scale_y_discrete(name = "Variance component") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```



# References