---
title: 'Supplementary Material for "Morph-dependent effect of nematode infection on host movement in the land snail *Cepaea nemoralis* (Mollusca, Gastropoda)"'
author: "Maxime Dahirel, Marine Proux, Claudia Gérard, Armelle Ansart"
date:
header-includes:
  - \usepackage{amsmath}
  - \usepackage{caption}
  - \DeclareCaptionLabelFormat{special}{#1#2}
output: pdf_document
bibliography: references.bib
csl: journal-of-zoology.csl
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE,
  message = FALSE, eval = TRUE, dev = "cairo_pdf"
)
```
<!--setting things up so that figures and tables captions are in the format "figure/table S[number]." since we are in the supplementary-->
<!--easy to adapt to journal requirements if needed-->
<!--reason for the use of caption in headers-includes-->
\captionsetup[table]{name=Table S,labelfont=bf, labelsep= period,labelformat=special}
\captionsetup[figure]{name=Figure S,labelfont=bf, labelsep= period,labelformat=special}

```{r packages-loading}
library(matrixStats) # CRAN v0.61.0

library(tidyverse) # CRAN v1.3.1
library(cmdstanr) # [github::stan-dev/cmdstanr] v0.4.0.9000
library(brms) # CRAN v2.16.3


library(bayesplot) # CRAN v1.8.1
library(tidybayes) # CRAN v3.0.2

library(patchwork) # CRAN v1.1.1

# for tables
library(knitr) # CRAN v1.37
library(kableExtra) # CRAN v1.3.4

library(here) # CRAN v1.0.1

options(mc.cores = 4) ## reduce if needed based on cores available
```


# S1 - Detailed model descriptions

In both models described below, we used weakly informative priors inspired by @mcelreathStatisticalRethinkingBayesian2020. We used $\mathrm{Normal}(0,1)$ priors for the fixed effects $\beta$, and $\mathrm{Half-Normal}(0,1)$ priors for all standard deviations (random effect standard deviations $\sigma_{\alpha}$ and $\sigma_{\gamma}$ as well as residual standard deviations $\sigma_{r}$). We used a $\mathrm{LKJ}(2)$ prior for the individual-level correlation matrix in the multivariate model.

## Body size (shell diameter) model

After centering and standardizing to unit 1SD, shell diameters $D_{o,m,i}$ with $o$ the population of origin (coded as a binary variable with 0 for the sun-exposed population and 1 for the shaded population), $m$ the shell morph and $i$ the individual, can be described by the following linear model:

$$
D_{o,m,i} \sim \mathrm{Normal}(\mu_{o,m},\sigma_{r[D]}), \\
$$
$$
\mu_{o,m} = \beta_{0[D,m]} + \beta_{1[D,m]} \times o. \\
$$
In this model, the intercepts $\beta_{0}$ are the morph-specific intercepts/mean sizes in the sun-exposed habitat, $\beta_{1}$ the morph-specific effect of the shaded habitat, and $\sigma_{r[D]}$ the residual variation.

## Multivariate model of parasite abundance and snail behaviours

The multivariate model used to analyse parasite abundances, movement behaviour, food intake and their individual-level correlations can be written as follows.

Let $N_{o,m,s,i}$ and $T_{o,m,s,i}$ be the numbers of live nematodes and *Brachylaima* trematodes (metacercariae) found in snail $i$ which was tested during session $s$, and $F_{o,m,s,i,t}$ and $M_{o,m,s,i,t}$ be the quantity of food ingested by snail $i$ (centered and scaled to 1SD), and its movement activity during trial $t$ (also centered and scaled to 1SD <!--COMM CLAUDIA: precise food consumption is proportion? ANSWER: no not necessary here, since for that model it does not matter-->), respectively. We then have:

$$
\begin{cases}
N_{o,m,s,i} \sim \mathrm{Poisson}(\lambda_{m,s,i[N]}), & \text{if}\ o = 0 \text{ (sun-exposed population)}\\
      N_{o,m,s,i} = 0, & \text{if}\ o = 1 \text{ (shaded population)},
\end{cases}
$$
$$
T_{o,m,s,i} \sim \mathrm{Poisson}(\lambda_{o,m,s,i[T]}),
$$
$$
F_{o,m,s,i,t} \sim \mathrm{Normal}(\mu_{o,m,s,i[F]},\sigma_{r[F]}),
$$
$$
M_{o,m,s,i,t} \sim \mathrm{Normal}(\mu_{o,m,s,i[M]},\sigma_{r[M]}),
$$
$$
\lambda_{m,s,i[N]} = \beta_{0[N,m]} + \beta_{2[N]} \times D_{i} + \alpha_{s[N]} + \gamma_{i[N]},\\
$$
$$
\lambda_{o,m,s,i[T]} = \beta_{0[T,m]} + \beta_{1[T,m]} \times o + \beta_{2[T]} \times D_{i} + \alpha_{s[T]} + \gamma_{i[T]},\\
$$
$$
\mu_{o,m,s,i[F]} = \beta_{0[F,m]} + \beta_{1[F,m]} \times o + \beta_{2[F]} \times D_{i} + \alpha_{s[F]} + \gamma_{i[F]},\\
$$
$$
\mu_{o,m,s,i[M]} = \beta_{0[M,m]} + \beta_{1[M,m]} \times o + \beta_{2[M]} \times D_{i} + \alpha_{s[M]} + \gamma_{i[M]},\\
$$
where the $\alpha$ refer to the experimental session random effects and $\gamma$ the individual random effects. These random effects are distributed as follows:
$$
\alpha_{s[N]} \sim \mathrm{Normal}(0,\sigma_{\alpha[N]}),
$$
$$
\alpha_{s[T]} \sim \mathrm{Normal}(0,\sigma_{\alpha[T]}),
$$
$$
\alpha_{s[F]} \sim \mathrm{Normal}(0,\sigma_{\alpha[F]}),
$$
$$
\alpha_{s[M]} \sim \mathrm{Normal}(0,\sigma_{\alpha[M]}),
$$
$$
\begin{bmatrix} \gamma_{i[N]} \\ \gamma_{i[T]} \\ \gamma_{i[F]} \\ \gamma_{i[M]} \end{bmatrix} 
\sim 
\textrm{MVNormal}
\begin{pmatrix}
\begin{bmatrix} 0 \\ 0 \\ 0 \\ 0  \end{bmatrix},
\boldsymbol{\Omega}
\end{pmatrix},
$$
, where $\boldsymbol{\Omega}$ is the individual-level covariance matrix, which can be decomposed into its constituent standard deviations and correlation matrix $\boldsymbol{R}$ in this way:
$$
\boldsymbol{\Omega} = 
\begin{bmatrix}
\sigma_{\gamma[N]} & 0 & 0 & 0\\
0 & \sigma_{\gamma[T]} & 0 & 0\\
0 & 0 & \sigma_{\gamma[F]} & 0\\
0 & 0 & 0 & \sigma_{\gamma[M]}
\end{bmatrix}
\boldsymbol{R}
\begin{bmatrix}
\sigma_{\gamma[N]} & 0 & 0 & 0\\
0 & \sigma_{\gamma[T]} & 0 & 0\\
0 & 0 & \sigma_{\gamma[F]} & 0\\
0 & 0 & 0 & \sigma_{\gamma[M]}
\end{bmatrix}.
$$

# S2 - Summary statistics on parasite infections and encapsulations

```{r data-load}
## we load the data needed as in the main script cepaea_2019_script.Rmd (see there for details)
raw_parasite <- read_csv(here("data", "cepaea_2019_parasite.csv"))
raw_behaviour <- read_csv(here("data", "cepaea_2019_behaviour.csv"))

data_parasite <- raw_parasite %>%
  mutate(population = fct_relevel(factor(population), "shaded", after = Inf)) %>% # factor reordering
  mutate(
    Nematode_live_all = Nematode_0_1_free + Nematode_1_3_free +
      Nematode_3_5_free + Nematode_5_plus_free + Nematode_1_3_lung +
      Nematode_3_5_lung + Nematode_5_plus_lung,
    Nematode_encaps_all = Nematode_white_encaps + Nematode_dark_encaps,
    Acari_live_all = Acari_kidney + Acari_digestive
  ) %>%
  mutate(session = 1)

data <- raw_behaviour %>%
  mutate(population = fct_relevel(factor(population), "shaded", after = Inf)) %>%
  left_join(data_parasite) %>%
  filter(dead == 0)

data <- data %>%
  # we then standardize size and response variables
  mutate(
    scale_size = scale(shell_diameter)[, 1],
    scale_food = scale(Food_intake)[, 1]
  ) %>%
  mutate(
    mean_size = mean(shell_diameter, na.rm = TRUE), ## useful for back-transformations for plots
    sd_size = sd(shell_diameter, na.rm = TRUE)
  ) %>%
  mutate(
    mean_food = mean(Food_intake, na.rm = TRUE),
    sd_food = sd(Food_intake, na.rm = TRUE)
  )
```

In the following tables, we present detailed descriptive information about the prevalence of each parasite (proportion of snails infected) and the infection intensity (number of parasites per *actually infected* snail). Parasite abundance, which we used as a synthetic measure of infection in the main text, is the mean number of parasites per snail *all snails included*; mean parasite abundance is equal to prevalence $\times$ mean intensity. See @bushParasitologyMeetsEcology1997 for details on definitions.  

For trematode and nematode data, the provided 95% intervals are Highest Density intervals based on Bernoulli (prevalence) and negative binomial (intensity) GLMs including shell morph, population and their interaction as fixed effects. The intensity models are fitted on (number of parasites - 1), to account for the zero truncation inherent to intensity. Priors are as in **S1**, with the addition of a $\mathrm{Half-Normal}(0,1)$ prior on the inverse of the negative binomial shape parameter, inspired by the authors of the Stan wiki: https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations.  

Note that we do not provide 95% intervals for mites data, given how rare infections are in our dataset. We also do not provide either intensity information or 95% intervals for encapsulated parasites, for reasons outlined at the end of the main text **Methods**. That is, contrary to *current* infections, the probability of having at least one parasite trapped and the total number of parasites trapped in the shell both increase cumulatively with age. Therefore, in the absence of information about age, there is no useful quantitative inference that can be made from these data. Note that encapsulated parasites were nonetheless counted and that these counts are made available along the rest of the raw data (see **Data availability** in main text). 

The tables below all exclude the individual that died during the experiment. That snail harboured `r subset(data_parasite$Nematode_live_all,data_parasite$dead==1)[1]` live nematodes, `r subset(data_parasite$Trematode_kidney,data_parasite$dead==1)[1]` live trematodes, `r subset(data_parasite$Acari_live_all,data_parasite$dead==1)[1]` live mites, had encapsulated `r subset(data_parasite$Nematode_encaps_all,data_parasite$dead==1)[1]` nematodes and `r subset(data_parasite$Acari_encaps,data_parasite$dead==1)[1]` mites.

```{r nematode-prevalence-model}
if (file.exists(here("R_output", "suppl_model_nema_prevalence.RDS"))) {
  mod_nemaprev <- readRDS(here("R_output", "suppl_model_nema_prevalence.RDS"))
} else {
  mod_nemaprev <- brm(
    bf(Nematode_live_all > 0 ~ 0 + band_number,
      family = bernoulli
    ),
    data = subset(data_parasite, population != "shaded"),
    chains = 4, iter = 2000, warmup = 1000,
    prior = c(
      set_prior("normal(0,1)", class = "b")
    ),
    backend = "cmdstanr", seed = 42
  )

  saveRDS(mod_nemaprev,
    file = here("R_output", "suppl_model_nema_prevalence.RDS")
  )
}
```

```{r nematode-intensity-model}
if (file.exists(here("R_output", "suppl_model_nema_intensity.RDS"))) {
  mod_nemaintens <- readRDS(here("R_output", "suppl_model_nema_intensity.RDS"))
} else {
  mod_nemaintens <- brm(
    bf(Nematode_live_all - 1 ~ 0 + band_number,
      nlf(shape ~ 1 / invshape),
      invshape ~ 1,
      family = negbinomial(link_shape = "identity")
    ),
    data = subset(data_parasite, Nematode_live_all > 0 & population != "shaded"),
    chains = 4, iter = 2000, warmup = 1000,
    prior = c(
      set_prior("normal(0,1)", class = "b"),
      set_prior("normal(0,1)", nlpar = "invshape", lb = 0)
    ),
    backend = "cmdstanr", seed = 42
  )

  saveRDS(mod_nemaintens,
    file = here("R_output", "suppl_model_nema_intensity.RDS")
  )
}
```

```{r trematode-prevalence-model}
if (file.exists(here("R_output", "suppl_model_trema_prevalence.RDS"))) {
  mod_tremaprev <- readRDS(here("R_output", "suppl_model_trema_prevalence.RDS"))
} else {
  mod_tremaprev <- brm(
    bf(Trematode_kidney > 0 ~ 0 + band_number + band_number:population,
      family = bernoulli
    ),
    data = data_parasite,
    chains = 4, iter = 2000, warmup = 1000,
    prior = c(
      set_prior("normal(0,1)", class = "b")
    ),
    backend = "cmdstanr", seed = 42
  )

  saveRDS(mod_tremaprev,
    file = here("R_output", "suppl_model_trema_prevalence.RDS")
  )
}
```

```{r trematode-intensity-model}
if (file.exists(here("R_output", "suppl_model_trema_intensity.RDS"))) {
  mod_tremaintens <- readRDS(here("R_output", "suppl_model_trema_intensity.RDS"))
} else {
  mod_tremaintens <- brm(
    bf(Trematode_kidney - 1 ~ 0 + band_number + band_number:population,
      nlf(shape ~ 1 / invshape),
      invshape ~ 1,
      family = negbinomial(link_shape = "identity")
    ),
    data = subset(data_parasite, Trematode_kidney > 0),
    chains = 4, iter = 2000, warmup = 1000,
    prior = c(
      set_prior("normal(0,1)", class = "b"),
      set_prior("normal(0,1)", nlpar = "invshape", lb = 0)
    ),
    backend = "cmdstanr", seed = 42
  )

  saveRDS(mod_tremaintens,
    file = here("R_output", "suppl_model_trema_intensity.RDS")
  )
}
```

```{r make-nematodes-tables}

# we prepare the summary statistics based directly on observed data
# like mean, N, min, max
nema_obs <- data_parasite %>%
  filter(dead == 0) %>%
  group_by(population, band_number) %>%
  summarise(
    n = length(Nematode_live_all),
    n_infected = length(subset(Nematode_live_all, Nematode_live_all > 0)),
    prevalence = mean(Nematode_live_all > 0) %>% round(digits = 2),
    mean_intensity = mean(subset(Nematode_live_all, Nematode_live_all > 0)) %>% round(digits = 2) %>% format(nsmall = 2),
    min_intensity = min(subset(Nematode_live_all, Nematode_live_all > 0)),
    max_intensity = max(subset(Nematode_live_all, Nematode_live_all > 0))
  ) %>%
  ungroup() %>%
  mutate(
    mean_intensity = replace(mean_intensity, which(mean_intensity == "NaN"), "--"),
    max_intensity = replace(max_intensity, which(max_intensity < 0), "--"),
    min_intensity = replace(min_intensity, which(min_intensity > 1000), "--"),
  )

# we use model predictions to create credible intervals

nema_pred_prev <- data_parasite %>%
  filter(population != "shaded") %>%
  select(band_number, population) %>%
  distinct() %>%
  add_epred_draws(mod_nemaprev) %>%
  mutate(prevalence = .epred) %>%
  group_by(band_number, population) %>%
  mean_hdi(prevalence) %>%
  group_by(band_number, population) %>%
  summarise(.lower = min(.lower), .upper = max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>%
  mutate(prevalence_interval = paste0("[", format(round(.lower, 2), nsmall = 2), "; ",
    format(round(.upper, 2), nsmall = 2), "]",
    sep = ""
  )) %>%
  select(band_number, population, prevalence_interval)


nema_pred_intens <- data_parasite %>%
  filter(population != "shaded") %>%
  select(band_number, population) %>%
  distinct() %>%
  add_epred_draws(mod_nemaintens) %>%
  mutate(intensity = .epred + 1) %>%
  group_by(band_number, population) %>%
  mean_hdi(intensity) %>%
  group_by(band_number, population) %>%
  summarise(.lower = min(.lower), .upper = max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>%
  mutate(intensity_interval = paste0("[", format(round(.lower, 2), nsmall = 2), "; ",
    format(round(.upper, 2), nsmall = 2), "]",
    sep = ""
  )) %>%
  select(band_number, population, intensity_interval)

# then we merge both

table_nema <- left_join(nema_obs, nema_pred_prev) %>%
  left_join(nema_pred_intens) %>%
  mutate(
    prevalence_interval = replace_na(prevalence_interval, "--"),
    intensity_interval = replace_na(intensity_interval, "--")
  ) %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  ) %>%
  select(
    population, band_number,
    n, n_infected, prevalence, prevalence_interval,
    mean_intensity, intensity_interval,
    min_intensity, max_intensity
  )
```

```{r make-and-display-trematodes-tables}
# we prepare the summary statistics based directly on observed data
# like mean, N, min, max
trema_obs <- data_parasite %>%
  filter(dead == 0) %>%
  group_by(population, band_number) %>%
  summarise(
    n = length(Trematode_kidney),
    n_infected = length(subset(Trematode_kidney, Trematode_kidney > 0)),
    prevalence = mean(Trematode_kidney > 0) %>% round(digits = 2),
    mean_intensity = mean(subset(Trematode_kidney, Trematode_kidney > 0)) %>% round(digits = 2) %>% format(nsmall = 2),
    min_intensity = min(subset(Trematode_kidney, Trematode_kidney > 0)),
    max_intensity = max(subset(Trematode_kidney, Trematode_kidney > 0))
  ) %>%
  ungroup()

# we use model predictions to create credible intervals


trema_pred_prev <- data_parasite %>%
  select(band_number, population) %>%
  distinct() %>%
  add_epred_draws(mod_tremaprev) %>%
  mutate(prevalence = .epred) %>%
  group_by(band_number, population) %>%
  mean_hdi(prevalence) %>%
  group_by(band_number, population) %>%
  summarise(.lower = min(.lower), .upper = max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>%
  mutate(prevalence_interval = paste0("[", format(round(.lower, 2), nsmall = 2), "; ",
    format(round(.upper, 2), nsmall = 2), "]",
    sep = ""
  )) %>%
  select(band_number, population, prevalence_interval)


trema_pred_intens <- data_parasite %>%
  select(band_number, population) %>%
  distinct() %>%
  add_epred_draws(mod_tremaintens) %>%
  mutate(intensity = .epred + 1) %>%
  group_by(band_number, population) %>%
  mean_hdi(intensity) %>%
  group_by(band_number, population) %>%
  summarise(.lower = min(.lower), .upper = max(.upper)) %>% ## to work around HDIs in two pieces
  ungroup() %>%
  mutate(intensity_interval = paste0("[", format(round(.lower, 2), nsmall = 2), "; ",
    format(round(.upper, 2), nsmall = 2), "]",
    sep = ""
  )) %>%
  select(band_number, population, intensity_interval)

# then we merge both

table_trema <- left_join(trema_obs, trema_pred_prev) %>%
  left_join(trema_pred_intens) %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  ) %>%
  select(
    population, band_number,
    n, n_infected, prevalence, prevalence_interval,
    mean_intensity, intensity_interval,
    min_intensity, max_intensity
  )
```

```{r make-mite-tables}
## mite table is observation data only

Acari_obs <- data_parasite %>%
  filter(dead == 0) %>%
  group_by(population, band_number) %>%
  summarise(
    n = length(Acari_live_all),
    n_infected = length(subset(Acari_live_all, Acari_live_all > 0)),
    prevalence = mean(Acari_live_all > 0) %>% round(digits = 2),
    mean_intensity = mean(subset(Acari_live_all, Acari_live_all > 0)) %>% round(digits = 2) %>% format(nsmall = 2),
    min_intensity = min(subset(Acari_live_all, Acari_live_all > 0)),
    max_intensity = max(subset(Acari_live_all, Acari_live_all > 0))
  ) %>%
  ungroup() %>%
  mutate(
    mean_intensity = replace(mean_intensity, which(mean_intensity == "NaN"), "--"),
    max_intensity = replace(max_intensity, which(max_intensity < 0), "--"),
    min_intensity = replace(min_intensity, which(min_intensity > 1000), "--"),
  ) %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )
```

```{r make-encaps-tables}
## encapsulations table is observation data only

encaps_obs <- data_parasite %>%
  filter(dead == 0) %>%
  group_by(population, band_number) %>%
  summarise(
    n = length(Acari_encaps),
    p_encaps_nema = mean(Nematode_encaps_all > 0) %>% round(digits = 2) %>% format(nsmall = 2),
    p_encaps_mite = mean(Acari_encaps > 0) %>% round(digits = 2) %>% format(nsmall = 2),
  ) %>%
  ungroup() %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )
```

```{r display-nematode-prevalence}
kable(table_nema %>%
  arrange(population, band_number) %>%
  select(
    `shell morph` = band_number, n,
    prevalence, `95% interval` = prevalence_interval
  ),
align = c("l","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about nematode prevalence"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)

### note you can add %>% kable_styling(latex_options = "HOLD_position")
### at the end of each table display snippet if the section break below (\newpage)
### isn't enough to keep them confined in place
```

```{r display-nematode-intensity}
kable(table_nema %>%
  arrange(population, band_number) %>%
  select(
    `shell morph` = band_number,
    `n(infected)` = n_infected,
    `mean intensity` = mean_intensity,
    `95% interval` = intensity_interval,
    min = min_intensity, max = max_intensity
  ),
align = c("l","r","r","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about nematode intensity of infection"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

```{r display-trematode-prevalence}
kable(table_trema %>%
  arrange(population, band_number) %>%
  select(
    `shell morph` = band_number, n,
    prevalence, `95% interval` = prevalence_interval
  ),
align = c("l","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about trematode prevalence"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

```{r display-trematode-intensity}
kable(table_trema %>%
  arrange(population, band_number) %>%
  select(
    `shell morph` = band_number,
    `n(infected)` = n_infected,
    `mean intensity` = mean_intensity,
    `95% interval` = intensity_interval,
    min = min_intensity, max = max_intensity
  ),
align = c("l","r","r","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about trematode intensity of infection"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

```{r display-mite-prevalence}
kable(Acari_obs %>%
  arrange(population, band_number) %>%
  select(`shell morph` = band_number, n, prevalence),
align = c("l","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about mite prevalence"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

```{r display-mite-intensity}
kable(Acari_obs %>%
  arrange(population, band_number) %>%
  select(
    `shell morph` = band_number,
    `n(infected)` = n_infected,
    `mean intensity` = mean_intensity,
    min = min_intensity, max = max_intensity
  ),
align = c("l","r","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about mite intensity of infection"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

```{r display-encaps}
kable(encaps_obs %>%
  arrange(population, band_number) %>% select(
    `shell morph` = band_number,
    n, `proportion with encapsulated nematodes` = p_encaps_nema,
    `proportion with encapsulated mites` = p_encaps_mite
  ),
align = c("l","r","r","r"),
booktabs = T,
caption = "Morph- and population-dependent summary information about parasite encapsulation in shells"
) %>%
  pack_rows("Open habitat", 1, 3) %>%
  pack_rows("Shaded habitat", 4, 6)
```

\newpage

# S3 - Pairwise posterior comparisons (multivariate model)

```{r loading-model}
mod_multi <- readRDS(here("R_output", "model_multi.RDS"))
```


```{r generate-predictions}
# " first we make predictions for each trait, much like in the main script
preds_food <- data %>%
  select(population, band_number, mean_food, sd_food) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1) %>%
  add_epred_draws(mod_multi, re_formula = NA, resp = "scalefood") %>%
  mutate(.epred = .epred * sd_food + mean_food) %>%
  ungroup() %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )

preds_mvt <- data %>%
  select(population, band_number) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1, is_active = TRUE) %>%
  add_epred_draws(mod_multi, re_formula = NA, resp = "scalemvt") %>%
  ungroup() %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )

preds_trematode <- data %>%
  select(band_number, population) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1) %>%
  add_linpred_draws(mod_multi, re_formula = NA, resp = "Trematodekidney") %>%
  ungroup() %>%
  mutate(
    population = fct_recode(population,
      `shaded habitat` = "shaded",
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )

preds_nematode <- data %>%
  select(band_number) %>%
  distinct() %>%
  mutate(scale_size = 0, session = 1, population = "sun_exposed") %>% ## predictions are made at the average size
  add_linpred_draws(mod_multi, re_formula = NA, resp = "Nematodeliveall") %>%
  ungroup() %>%
  mutate(
    population = fct_recode(population,
      `open habitat` = "sun_exposed"
    ),
    band_number = fct_recode(band_number,
      `0 bands` = "0B",
      `3 bands` = "3B",
      `5 bands` = "5B"
    )
  )
```

```{r fig.cap="Posterior pairwise differences in nematode abundance between snail morphs (on the latent log scale). Reminder: the model for nematode abundance only includes data from the sun-exposed population, since no live nematodes were found in the shaded population.", fig.height=3}
# then we do pairwise comparisons and plot them for each trait (here nematode abundance)
preds_nematode %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.linpred, by = group) %>%
  ggplot() +
  stat_halfeye(aes(.linpred, group),
    .width = c(0.001, 0.95), point_interval = "mean_hdi"
  ) +
  geom_vline(xintercept = 0, lty = 2) +
  scale_y_discrete("pair compared") +
  scale_x_continuous("posterior predicted difference (latent log scale)") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```


```{r fig.cap="Posterior pairwise differences in *Brachylaima* trematode (metacercariae) abundance between snail morph × population combinations (on the latent log scale).", fig.height=6}
preds_trematode %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.linpred, by = group) %>%
  ggplot() +
  stat_halfeye(aes(.linpred, group),
    .width = c(0.001, 0.95), point_interval = "mean_hdi"
  ) +
  geom_vline(xintercept = 0, lty = 2) +
  scale_y_discrete("pair compared") +
  scale_x_continuous("posterior predicted difference (latent log scale)") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```


```{r fig.cap="Posterior pairwise differences in movement behaviour between snail morph × population combinations. The movement variable is the inverse of the latency to move away (so higher values mean more mobile individuals), and the variable was scaled so differences are in units 1SD.", fig.height=6}
preds_mvt %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.epred, by = group) %>%
  ggplot() +
  stat_halfeye(aes(.epred, group),
    .width = c(0.001, 0.95), point_interval = "mean_hdi"
  ) +
  geom_vline(xintercept = 0, lty = 2) +
  scale_y_discrete("pair compared") +
  scale_x_continuous("posterior predicted difference") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```

```{r fig.cap="Posterior pairwise differences in food intake between snail morph × population combinations.", fig.height=6}
preds_food %>%
  mutate(group = paste(band_number, ", ", population, sep = "")) %>%
  compare_levels(.epred, by = group) %>%
  ggplot() +
  stat_halfeye(aes(.epred, group),
    .width = c(0.001, 0.95), point_interval = "mean_hdi"
  ) +
  geom_vline(xintercept = 0, lty = 2) +
  scale_y_discrete("pair compared") +
  scale_x_continuous("posterior predicted difference") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```

\newpage

# S4 - Posterior proportion of total variance associated with fixed effects vs. random effects in movement and food intake

Both movement behaviour and food intake were observed twice, which allows us to partition variance into among-individual and within-individual components. We show that although in both cases, within-individual/"residual" variation is the dominant component, there is a non-negligible among-individual variance component: both traits are repeatable to some extent (**Supporting Information Figure S5**).

```{r prep-fig-varpart}
## the variables used to fit models are scaled
## but that's not a problem since we are plotting **relative** variance components here

## in each case there are four variance components:
## the fixed effects, the individual-level random effects,
## the experimental session random effects, and the residual within-individual variation


## here for the movement model
varcomps_mvt <- tibble(
  VF = rowVars(posterior_linpred(mod_multi, re_formula = NA, resp = "scalemvt")),
  VI = VarCorr(mod_multi, summary = FALSE)$fullID$sd[, "scalemvt_Intercept"]^2,
  Vseries = VarCorr(mod_multi, summary = FALSE)$series$sd[, "scalemvt_Intercept"]^2,
  VR = as_tibble(as_draws_df(mod_multi, variable = "sigma_scalemvt"))$sigma_scalemvt^2
) %>%
  mutate(
    trait = "mvt",
    VP = VF + VI + Vseries + VR
  )

## and here for the food intake model
varcomps_food <- tibble(
  VF = rowVars(posterior_linpred(mod_multi, re_formula = NA, resp = "scalefood")),
  VI = VarCorr(mod_multi, summary = FALSE)$fullID$sd[, "scalefood_Intercept"]^2,
  Vseries = VarCorr(mod_multi, summary = FALSE)$series$sd[, "scalefood_Intercept"]^2,
  VR = as_tibble(as_draws_df(mod_multi, variable = "sigma_scalefood"))$sigma_scalefood^2
) %>%
  mutate(
    trait = "food",
    VP = VF + VI + Vseries + VR
  )

## we merge them and do some rearranging/ renaming
varcomps <- rbind(varcomps_food, varcomps_mvt) %>%
  group_by(trait) %>%
  pivot_longer(
    cols = c(VF, VI, Vseries, VR),
    names_to = "varcomp", values_to = "variance"
  ) %>%
  ungroup() %>%
  mutate(varcomp = fct_recode(varcomp,
    `fixed effects` = "VF",
    individual = "VI",
    `test group` = "Vseries",
    residuals = "VR"
  )) %>%
  mutate(varcomp = fct_relevel(varcomp, rev)) %>%
  mutate(varcomp = fct_relevel(varcomp, "test group", after = 1)) %>%
  mutate(trait = fct_recode(trait,
    `Food intake` = "food",
    `Movement (1/latency to leave)` = "mvt"
  ))
```


```{r fig-varpart, fig.cap="Mean (points) and posteriors for the proportion of behavioural variance explained by the different variance components. See **Methods** and **Supporting Information S1** for a description of the model underlying these estimates.", fig.height= 3}

ggplot(varcomps) +
  stat_halfeye(aes(y = varcomp, x = 100 * variance / VP),
    point_interval = mean_hdi, .width = c(0.001, 0.95),
    normalize = "xy"
  ) +
  facet_wrap(~trait) +
  scale_x_continuous(
    name = "Variance explained (%, observed scale)", lim = c(0, 100),
    breaks = seq(from = 0, to = 100, by = 10)
  ) +
  scale_y_discrete(name = "Variance component") +
  cowplot::theme_half_open(11) +
  cowplot::background_grid(colour.major = "grey95", colour.minor = "grey95")
```

# References
